---
title: "Reproducible research"
author: "Dan Taylor Lewis"
date: "07/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T)
```


### Read in the data and set librarys.

```{r,message=FALSE}
library(dplyr)
library(ggplot2)
```

```{r,message=F,results='hide',warning=F}
dir.create("C:/Users/Dan/Documents/Coursera/Repoducible Research/")
setwd("C:/Users/Dan/Documents/Coursera/Repoducible Research/")
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip","activities monitoring data.zip")
list.files()
unzip("activities monitoring data.zip")
activity<-read.csv("activity.csv")
```

#### Quick look at data set date as a date.

```{r,results='hide'}
lapply(activity, function(x) summary(x))
activity$date<-as.Date(activity$date)

```

### Histogram of total number of steps taken each day.

```{r}
activity%>%filter(!is.na(steps))%>%group_by(date)%>%summarise(total_steps=sum(steps))%>%
    ggplot(aes(x=total_steps))+geom_histogram(binwidth = 1000)+labs(title="Histogram: Total steps taken each day",x="Total Steps",y="Count")
```
### Mean and median number of steps taken each day
```{r}
total_steps_by_day<-activity%>%group_by(date)%>%summarise(total_steps=sum(steps))
mean(total_steps_by_day$total_steps,na.rm=T)
median(total_steps_by_day$total_steps,na.rm=T)
```

### Time series plot of the average number of steps taken

```{r}
activity%>%filter(!is.na(steps))%>%group_by(date)%>%summarise(average_steps_per_interval=mean(steps))%>%
    ggplot(aes(x=date,y=average_steps_per_interval,group=1))+geom_point()+geom_line(stat="identity",color="blue")+labs(title = "Time series: Average steps per interval by day",x="Date",y="Average step count")+theme(axis.text.x = element_text(angle = 90))

```

### The 5-minute interval that, on average, contains the maximum number of steps

```{r}
activity%>%filter(!is.na(steps))%>%group_by(interval)%>%summarise(average_steps=mean(steps))%>%arrange(desc(average_steps))%>%head(.,1)
```

### Code to describe and show a strategy for imputing missing data


I will compute the average steps using the days and intervals without missing data and then impute the missing intervals to the corresponding average

```{r}
impute_to<-activity%>%filter(!is.na(steps))%>%group_by(interval)%>%summarise(avg_steps=mean(steps))
for (i in 1:dim(activity)[1]){
    if (is.na(activity[i,"steps"])){
        activity[i,"steps"]<-impute_to[which(impute_to[,"interval"]==activity[i,"interval"]),"avg_steps"]
    }

}

#Check 

sum(is.na(activity$steps))

```

### Histogram of total number of steps taken each day after imputation.

```{r}
activity%>%group_by(date)%>%summarise(total_steps=sum(steps))%>%
    ggplot(aes(x=total_steps))+geom_histogram(binwidth = 1000)+labs(title="Histogram: Total steps taken each day",x="Total Steps",y="Count")
```

### Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}

activity$day<-as.factor(weekdays(activity$date))
levels(activity$day)
activity$weekend_flag<-as.factor(ifelse(as.numeric(activity$day) %in% c(3,4),1,0))
levels(activity$weekend_flag)<-c("Weekday", "Weekend")


activity%>%group_by(interval,weekend_flag)%>%summarise(avg_steps=mean(steps))%>%
    ggplot(.,aes(x=interval,y=avg_steps,group=weekend_flag))+geom_point()+geom_line(color="blue")+ facet_wrap(~ weekend_flag)+labs(x="Interval",y="Average Step Count", title = "Average steps per 5 minute interval across weekdays and weekends")
    
 
```


